---
title: README
output: github_document
---

The code is designed to be run using the ```drake``` package, by running the 
code in the script [make.R](make.R) either via the command line 
(i.e., ```R CMD BATCH make.R```), or from within an interactive ```R``` session.
The outline of all project components (i.e., the ```drake``` plan) is 
available in the [R/subplans](R/subplans) directory.

The ```drake``` [Github page](https://github.com/ropensci/drake) page has a 
good overview of what drake does, and includes some code snippets.  After 
running the [make.R](make.R) script, the ```drake::loadd``` function is the 
most helpful function for loading output.

The ```drake``` [documentation](https://books.ropensci.org/drake/index.html) 
itself is also a good resource for learning how to use the ```drake``` package.  
Starting with the 
[Walkthrough](https://books.ropensci.org/drake/walkthrough.html#set-the-stage.)
is a decent place to just jump in to some details.


# Data formats

The Error measurement model is implemented in the 
[model_joint.R](R/nimble/model_joint.R) script.  The model implementation is 
abstract, to allow multiple observations of any object.  The 
[data_plan.R](R/subplans/data_plan.R) script creates the necessary, raw data 
structures for the model.  The [flatten_data()](R/nimble/flatten_data.R)
function munges the data structures for use with the model's implementation
in ```nimble```.


## Training object information

Each training object must be documented in the same format as the 
```training_obj``` object (below).  The first two columns specify the 
training object, and the third column records its known length.  The 
[flatten_data()](R/nimble/flatten_data.R)
function will use the information in ```training_obj``` to 
associate known lengths with image and pixel measurements of the training 
objects.


```{r, echo = FALSE}
library(drake)
loadd(training_obj)
knitr::kable(training_obj, row.names = FALSE)
```


## Image data

Each image must be documented in the same format as the ```Mns_images``` object
(preview below).  The first column ```Image``` provides a unique identifier for 
the image and will be used to create an altitude variable, which will be 
estimated from the altitude sensor readings provided by ```AltitudeBarometer``` 
and ```AltitudeLaser``` in conjunction with training data.  The remaining 
columns in ```Mns_images``` report the focal length (mm), image width (pixels), 
and camera sensor width (mm) associated with the image, which will be used to 
estimate the image's ground sampling distance (GSD) per pixel.

```{r, echo = FALSE}
library(drake)
loadd(Mns_images)
knitr::kable(Mns_images[1:5,])
```

## Pixel count "measurement" data

All pixel measurements must also be documented in the same format as the 
```Mns_pixels``` object (preview below).  The first two columns specify the 
measurement.  One measurement variable, which will be estimated, will be 
associated with each unique combination of the first two columns. The 
```Image``` column links the measurement to an image and it's estimated GSD.
Lastly, the ```PixelCount``` column records the pixel-length of the object as 
it appears in the image.  

```{r, echo = FALSE}
library(drake)
loadd(Mns_pixels)
knitr::kable(Mns_pixels[3:8,])
```

## Many-to-Many relationships

Storing Image and Pixel count data in separate structures allows multiple 
measurements to be estimated from a single image.  For example, total length and 
maximum width can be measured from the same image through a pixel count table 
like the following:

```{r, echo = FALSE}
knitr::kable(
  data.frame(Subject = 'Animal A', Measurement = c('Total length', 'Max. width'), 
           Image = 'Image 1', PixelCount = c(1e3, 1e2))
)

```

Both ```PixelCount``` entries above relate to a different object being measured, 
but use the same estimated altitude and implied GSD for ```Image 1``` when
estimating lengths from pixel counts.


Similarly, the Image and Pixel count structures also allow one object to be 
estimated from multiple images.  For example, total length can be estimated 
from multiple observations through a pixel count table like the following:

```{r, echo = FALSE}
knitr::kable(
  data.frame(Subject = 'Animal A', Measurement = 'Total length', 
           Image = paste('Image', 1:5, sep = ' '), 
           PixelCount = c(1e3, rnorm(n = 4, mean = 1e3, sd = 20)))
)
```



## Materials

### Data:

"APE-Dataset.csv" - training/calibration data for different UAS platforms.

Images = image used for measuring calibration object
Aircraft = the UAS aircraft used to collect imagery of calibration object to measure
Flight = the flight number/name
Position = indicates if calibration object is in center of corner of image frame
Focal.length = focal length of camera
Pix.Dim = pixel dimensions (mm/px), which is the sensor width (mm) / image width (px) of the camera
Baro_raw = the raw relative altitude recorded by the barometer
Launch.Ht = the launch height of the drone, to be added to the BarAlt to get the absolute barometric altitude above the surface of the water
Baro + Ht = the baro_raw + Launch.Ht to get the absolute barometer altitude
Laser_Alt = the altitude recorded by the laser altimeter. Blanks spaces/NAs indicate no/false reading
Alitude = altitude used in measurement, either Laser_Alt or Baro + Ht
Alimeter = which altimeter was used for altitude in measurement; either barometer or laser
Measurer = which analyst measured the calibration object (3 analysts total, measured each image once)
RRR.pix = the length in pixels of the known sized object (calibration object)  


*"humpback_data.csv" - testing data. Humpback lengths from the Western Antarctic Peninsula.

+Animal_ID = unique ID for individual whale
+Reproductive_Class = determined from biopsy sample or from drone images (mom and calf pairs)
+Species = the species of whale
+Image = the image ID used for measuring the whale
+Aircraft = which UAS platform collected the imagery
+Focal_Length = focal length (mm) of the camera used
+LaserAlt = the altitude recorded by the laser altimeter. Blanks spaces/NAs indicate no/false reading
+BaroAlt = the raw relative altitude recorded by the barometer
+Pixel_Dimension = (mm/px) the Sensor width (mm) / Image width (px) of the camera
+Launch_Ht = the launch height of the drone, to be added to the BarAlt to get the absolute barometric altitude above the surface of the water
+Altitude = altitude used in measurement, either LaserAlt or BaroAlt + Launch_Ht
+Altimeter = which altimeter was used for altitude in measurement; either barometer or laser
+TL = total length of the whale in meters