---
title: README
output: github_document
---

This code is designed to be run using the ```drake``` package, by running the 
code in the script [make.R](make.R) either via the command line 
(i.e., ```R CMD BATCH make.R```), or from within an interactive ```R``` session.
The outline of all project components (i.e., the ```drake``` plan) is 
available in the [R/subplans](R/subplans) directory.

The ```drake``` [Github page](https://github.com/ropensci/drake) page has a 
good overview of what drake does, and includes some code snippets.  After 
running the [make.R](make.R) script, the ```drake::loadd``` function is the 
most helpful function for loading output.

The ```drake``` [documentation](https://books.ropensci.org/drake/index.html) 
itself is also a good resource for learning how to use the ```drake``` package.  
Starting with the 
[Walkthrough](https://books.ropensci.org/drake/walkthrough.html#set-the-stage.)
is a decent place to just jump in to some details.


&nbsp;
&nbsp;

Code and data from:    
**A Bayesian approach for predicting photogrammetric uncertainty in morphometric measurements derived from UAS**  
KC Bierlich, Robert Schick, Josh Hewitt, Julian Dale, Jeremy Goldbogen, Ari Friedlaender, David Johnston  
*Marine Ecology Progress Series*  
DOI: https://doi.org/10.3354/meps13814. 

# Data formats

The Error measurement model is implemented in the 
[model_joint.R](R/nimble/model_joint.R) script.  The model implementation is 
abstract, to allow multiple observations of any object.  The 
[data_plan.R](R/subplans/data_plan.R) script creates the necessary, raw data 
structures for the model.  The [flatten_data()](R/nimble/flatten_data.R)
function munges the data structures for use with the model's implementation
in ```nimble```.


## Training object information

Each training object must be documented in the same format as the 
```training_obj``` object (below).  The first two columns specify the 
training object, and the third column records its known length.  The 
[flatten_data()](R/nimble/flatten_data.R)
function will use the information in ```training_obj``` to 
associate known lengths with image and pixel measurements of the training 
objects.


```{r, echo = FALSE}
library(drake)
loadd(training_obj)
knitr::kable(training_obj, row.names = FALSE)
```


## Image data

Each image must be documented in the same format as the ```Mns_images``` object
(preview below).  The first column ```Image``` provides a unique identifier for 
the image and will be used to create an altitude variable, which will be 
estimated from the altitude sensor readings provided by ```AltitudeBarometer``` 
and ```AltitudeLaser``` in conjunction with training data.  The remaining 
columns in ```Mns_images``` report the focal length (mm), image width (pixels), 
and camera sensor width (mm) associated with the image, which will be used to 
estimate the image's ground sampling distance (GSD) per pixel.

```{r, echo = FALSE}
library(drake)
loadd(Mns_images)
knitr::kable(Mns_images[1:5,])
```

## Pixel count "measurement" data

All pixel measurements must also be documented in the same format as the 
```Mns_pixels``` object (preview below).  The first two columns specify the 
measurement.  One measurement variable, which will be estimated, will be 
associated with each unique combination of the first two columns. The 
```Image``` column links the measurement to an image and it's estimated GSD.
Lastly, the ```PixelCount``` column records the pixel-length of the object as 
it appears in the image.  

```{r, echo = FALSE}
library(drake)
loadd(Mns_pixels)
knitr::kable(Mns_pixels[3:8,])
```

## Many-to-Many relationships

Storing Image and Pixel count data in separate structures allows multiple 
measurements to be estimated from a single image.  For example, total length and 
maximum width can be measured from the same image through a pixel count table 
like the following:

```{r, echo = FALSE}
knitr::kable(
  data.frame(Subject = 'Animal A', Measurement = c('Total length', 'Max. width'), 
           Image = 'Image 1', PixelCount = c(1e3, 1e2))
)

```

Both ```PixelCount``` entries above relate to a different object being measured, 
but use the same estimated altitude and implied GSD for ```Image 1``` when
estimating lengths from pixel counts.


Similarly, the Image and Pixel count structures also allow one object to be 
estimated from multiple images.  For example, total length can be estimated 
from multiple observations through a pixel count table like the following:

```{r, echo = FALSE}
knitr::kable(
  data.frame(Subject = 'Animal A', Measurement = 'Total length', 
           Image = paste('Image', 1:5, sep = ' '), 
           PixelCount = c(1e3, rnorm(n = 4, mean = 1e3, sd = 20)))
)
```



## Materials

### Data:

_APE-Dataset.csv_ - training/calibration data for different UAS platforms.  
 * Images = image used for measuring calibration object  
 * Aircraft = the UAS aircraft used to collect imagery of calibration object to measure  
 * Flight = the flight number/name  
 * Position = indicates if calibration object is in center of corner of image frame  
 * Focal.length = focal length of camera  
 * Pix.Dim = pixel dimensions (mm/px), which is the sensor width (mm) / image width (px) of the camera  
 * Baro_raw = the raw relative altitude recorded by the barometer  
 * Launch.Ht = the launch height of the drone, to be added to the BarAlt to get the absolute barometric altitude above the surface of the water  
 * Baro + Ht = the baro_raw + Launch.Ht to get the absolute barometer altitude  
 * Laser_Alt = the altitude recorded by the laser altimeter. Blanks spaces/NAs indicate no/false reading  
 * Alitude = altitude used in measurement, either Laser_Alt or Baro + Ht  
 * Alimeter = which altimeter was used for altitude in measurement; either barometer or laser  
 * Measurer = which analyst measured the calibration object (3 analysts total, measured each image once)  
 * RRR.pix = the length in pixels of the known sized object (calibration object)    
  

_calibration_object-measurements.csv_ - training/calibration data for the LemHex-44 and FreeFly Alta 6 to add to both Ecological Scenarios  
  * CO.ID = calibration object ID  
  * CO.Length = the true length of the CO.ID  
  * Image = image used for measuring calibration object  
  * Lab = the lab that collected the data  
  * Cruise = research expedicition ID  
  * Date = date that imagery was collected of calbiration object  
  * Flight = the flight number/name  
  * Pilot = pilot during data collection  
  * VO = visual observer  
  * Aircraft = the UAS aircraft used to collect imagery of calibration object to measure  
  * Focal_length = focal length of camera  
  * Iw = image width in pixels  
  * Sw = sensor width in mm  
  * pix.dim = pixel dimensions; Sw/Iw
  * Pix.Dim = pixel dimensions (mm/px), which is the sensor width (mm) / image width (px) of the camera  
  * Baro_raw = the raw relative altitude recorded by the barometer  
  * Launch.Ht = the launch height of the drone, to be added to the BarAlt to get the absolute barometric altitude above the surface of the water  
  * Baro+Ht = the baro_raw + Launch.Ht to get the absolute barometer altitude  
  * Laser_Alt = the altitude recorded by the laser altimeter. Blanks spaces/NAs indicate no/false reading
  * Alitude = altitude used in measurement, either Laser_Alt or Baro + Ht  
  * Alimeter = which altimeter was used for altitude in measurement; either barometer or laser  
  * Lpix = the length in pixels of the known sized object (calibration object)    
  * object_position = indicates if calibration object is in center of corner of image frame  
  * Analyst = analyst that performed the measurement
 


_humpback_data.csv_ - testing data for Ecological Scenario 1, length-based maturity of humpback whales
 * whale = numbered list of individuals  
 * Animal_ID = unique ID for individual whale  
 * Reproductive_Class = determined from biopsy sample or from drone images (mom and calf pairs)  
 * Species = the species of whale  
 * Image = the image ID used for measuring the whale  
 * Aircraft = which UAS platform collected the imagery  
 * Focal_Length = focal length (mm) of the camera used  
 * LaserAlt = the altitude recorded by the laser altimeter. Blanks spaces/NAs indicate no/false reading  
 * BaroAlt = the raw relative altitude recorded by the barometer  
 * Pixel_Dimension = (mm/px) the Sensor width (mm) / Image width (px) of the camera  
 * Launch_Ht = the launch height of the drone, to be added to the BarAlt to get the absolute barometric altitude above the surface of the water  
 * Altitude = altitude used in measurement, either LaserAlt or BaroAlt + Launch_Ht  
 * Altimeter = which altimeter was used for altitude in measurement; either barometer or laser  
 * TL = total length of the whale in meters  


_marrs_amws.csv_ - testing data for Ecological Scenario 2, population-lelve morphological relationship between rostrum to blowhole distance and total body length
 * AID = unique ID for individual whale  
 * Species = the species of whale  
 * Image = the image ID used for measuring the whale  
 * Region = location of data collection (WAP = Western Antarctic Peninsula)  
 * Year = year of data collection  
 * Aircraft = which UAS platform collected the imagery  
 * TL = total length of the whale in meters  
 * RB = rostrum to blowhole distance in meters  
 * BaroAlt = the raw relative altitude recorded by the barometer  
 * Launch_Ht = the launch height of the drone, to be added to the BarAlt to get the absolute barometric altitude above the surface of the water  
 * LaserAlt = the altitude recorded by the laser altimeter. Blanks spaces/NAs indicate no/false reading  
 * Altitude = altitude used in measurement, either LaserAlt or BaroAlt + Launch_Ht  
 * Focal_Length = focal length (mm) of the camera used  
 * Iw = image width in pixels  
 * Sw = sensor width in mm  
 * perDif = percent difference between the altitude recorded by the barometer and laser altimeter  


## Steps

run 'make.R'

This will generate an "output" folder that contains an 'mcmc' folder and a 'reports' folder.   
 
  * The 'mcmc' folder contains all the .rds outputs from the model (note that "fit_mns" refers to humpbacks in Ecological Scenario 1, while "fit_marrs" refers to Antarctic minke whales in Ecological Scenario 2), and a 'diagnostics' folder with .html files providing diagnostic comparisons between the different model outputs (e.g., barometer only, barometer and laser, uninformative priors, etc.).   
  
  * The 'reports' folder contains .html files to evaluate the results from the model  
      + Sc1_output_figures explores outputs from Ecological Scenario 1. Figures are automatically saved in the "figures" folder of the main directory.
      
      + fit_marrs refers to results from Scenario 2  
      
      + Uncorrected_error_figures explores the APE dataset (training data) before applying the uncertainty model. Figures are automatically saved in the "figures" folder of the main directory.  
      
  * Figures 3-6 and Supplementary FigS1 and FigS2 are automatically generated in the 'figures' folders
  




## Contact
KC Bierlich, kevin.bierlich@oregonstate.edu  
https://github.com/KCBierlich  

