---
title: "Output Figures"
author: "KCB"
date: "7/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
library(dotenv)
library(drake)
library(coda)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(knitr)
library(nimble)
library(scoringRules)
library(stringr)
library(ggdist)
library(tidyr)
library("ggpubr")
library(conflicted) 
conflict_prefer("filter", "dplyr")
conflict_prefer("summarize", "dplyr")
```




# Ecological Scenario 1: length based maturity classification for humpback whales
Model 1 = barometer only  
Model 2 = barometer and laser  
  
load model outputs  
```{r}
setwd("/Users/kcbierlich/Dropbox/GitHub/Uncertainty_Model/reports")
mns_mod1 <- readRDS(file.path("..", "output", "mcmc","length_samples_mns_fit_mns_barometer.rds"))
mns_mod2 <- readRDS(file.path("..", "output", "mcmc","length_samples_mns_fit_mns_barometer.laser.rds"))


mod1_samples <- as.list(mns_mod1$samples)
mod2_samples <- as.list(mns_mod2$samples)

```


Compute probability of maturity
```{r}
#
# Model 1: Barometer only
#

full_df <-  mod1_samples       # save individual's dataframes to be used in the loop for maturity
ID_list <-  names(full_df)     # save list of AIDs     

# make an empty dataframe to fill with the summary stats
maturity.df <- data.frame()

# loop through list of AIDs (loop through each individual's data frame) and calculate the proportion of TLs > 11.2. 
# result has a column of 'perMature' representing the "percent likelihood of being mature"
for (x in 1:length(ID_list)){
  AID <- ID_list[x]       # pull ID name from list & save as variable
  xx <-  full_df[[AID]]   # save TL precictions 
  xx.t <- xx$`Total length` > 11.2       # set maturity cutoff and save values where TL is greater than cutoff 
  perMature <-  round((length(xx.t[xx.t == TRUE])/length(xx$`Total length`))*100,2)  # calculate %% TL prediction > 11.2
  mTL <- mean(xx$`Total length`)
  hpd <- HPDinterval(mcmc(xx$`Total length`))
  hpd_lwr <- hpd[1]
  hpd_upr <- hpd[2]

  # merge
  tempdf <- data.frame(AID, mTL, hpd_lwr, hpd_upr, perMature)  # save output in temp df
  maturity.df <- rbind(maturity.df,tempdf)   # add the row of summary stats for this individual to the main dataframe
}
  
mat.mod1 <- maturity.df # save maturity output for Model 1


#
# Model 2: Barometer and Laser
#

full_df <-  mod2_samples       # save individual's dataframes to be used in the loop for maturity
ID_list <-  names(full_df)     # save list of AIDs     

# make an empty dataframe to fill with the summary stats
maturity.df <- data.frame()

# loop through list of AIDs (loop through each individual's data frame) and calculate the proportion of TLs > 11.2. 
# result has a column of 'perMature' the "percent likelihood of being mature"
for (x in 1:length(ID_list)){
  AID <- ID_list[x]       # pull ID name from list & save as variable
  xx <-  full_df[[AID]]   # save TL predictions 
  xx.t <- xx$`Total length` > 11.2       # set maturity cutoff and save values where TL is greater than cutoff 
  perMature <-  round((length(xx.t[xx.t == TRUE])/length(xx$`Total length`))*100,2)  # calculate %% TL prediction > 11.2
  mTL <- mean(xx$`Total length`)
  hpd <- HPDinterval(mcmc(xx$`Total length`))
  hpd_lwr <- hpd[1]
  hpd_upr <- hpd[2]

  # merge
  tempdf <- data.frame(AID, mTL, hpd_lwr, hpd_upr, perMature)  # save output in temp df
  maturity.df <- rbind(maturity.df,tempdf)   # add the row of summary stats for this individual to the main dataframe
}
  
mat.mod2 <- maturity.df # save maturity output for Model 2

# these AIDs have NA barometer values and need to be omitted.
mat.mod1 <- mat.mod1 %>% dplyr::filter(AID != "Mn190325_A_F1-02")
mat.mod1 <- mat.mod1 %>% dplyr::filter(AID != "Mn190325_A_F1-01")
mat.mod2 <- mat.mod2 %>% dplyr::filter(AID != "Mn190325_A_F1-02")
mat.mod2 <- mat.mod2 %>% dplyr::filter(AID != "Mn190325_A_F1-01")

```



## Figure 4
Example of predictive posterior TL distribution for an individual whale
```{r}
## Find a good example of an individual with different classifications between on Model 1 and 2
  # mat.mod1 %>% dplyr::filter(perMature < 60 & perMature > 45)
  # mat.mod2 %>% dplyr::filter(perMature < 28 & perMature > 1)

Ex_mod1 <- mns_mod1$samples$"Mn190318_A_F1-01"
Ex_mod2 <- mns_mod2$samples$"Mn190318_A_F1-01"

# # Assign variable for Model
Ex_mod1$altimeter <-  "Model 1: barometer"
Ex_mod2$altimeter <- "Model 2: barometer & laser"

# Combine
Ex <- rbind(Ex_mod1, Ex_mod2)
names(Ex)[names(Ex) == "Total length"] <- "TL"

```


Plot Figure 4
```{r}
ggplot(Ex %>% dplyr::filter(TL < 15), aes(TL)) + 
  stat_halfeye() +
  xlab("predicted length (m)") +
  ylab("density") +
  facet_wrap(~altimeter, ncol = 1) + 
  geom_vline(aes(xintercept=11.2), col = "red", lty = 2, alpha = 0.75) 
ggsave(file.path('..', 'figures', 'Fig4.jpg'), width = 7, height = 7)
```


## Figure 5. 
Comparison of Model 1 and Model 2 outputs  
Humpback predicted lengths w/ increasing altitude & maturity classification
```{r}

mod_inputs <- read.csv(file.path("..", "data", "humpback_data.csv"))
names(mod_inputs)[names(mod_inputs) == "Animal_ID"] <- "AID"

# Merge metadata from model inputs 
mat.mod1 <- mat.mod1 %>% left_join(mod_inputs, by = "AID")
mat.mod2 <- mat.mod2 %>% left_join(mod_inputs, by = "AID")

## Predicted humpback lengths for Model 1 and 2
M1.output <- ggplot(mat.mod1, aes(x = BaroAlt, y = mTL, ymin = hpd_lwr, ymax = hpd_upr, color = perMature)) + 
  geom_pointrange() + #colour = "royalblue2") + 
  scale_x_continuous(breaks = seq(10, 80, by = 10)) +
  scale_y_continuous(breaks = seq(7, 18, by = 1)) + 
  scale_color_continuous(name = "% Maturity") +
  xlab('observed altitude (m)') + 
  ylab('predicted length (m)') +
  ggtitle("Model 1: barometer") +
  geom_abline(aes(intercept=11.2, slope=0), col = "red", lty = 2, alpha = 0.75)  # Cutoff for mature humpback whales (Christiansen et al., 2016)


M2.output <- ggplot(mat.mod2, aes(x = BaroAlt, y = mTL, ymin = hpd_lwr, ymax = hpd_upr, color = perMature)) + 
  geom_pointrange() + #colour = "royalblue2") + 
  scale_x_continuous(breaks = seq(10, 80, by = 10)) +
  scale_y_continuous(breaks = seq(7, 18, by = 1)) + 
  xlab('observed altitude (m)') + 
  ylab('predicted length (m)') +
  #ylim(7,18) +
  ggtitle("Model 2: barometer & laser") +
  scale_color_continuous(name = "% Maturity") +
  geom_abline(aes(intercept=11.2, slope=0), col = "red", lty = 2, alpha = 0.75)  # Cutoff for mature humpback whales (Christiansen et al., 2016) 

ggarrange(M1.output, M2.output, ncol = 1, nrow = 2)
ggsave(file.path('..', 'figures', 'Fig5.jpg'), width = 7, height = 7)
``` 



## Figure 6.
Comparing Model 1 and Model 2 measurement uncertainty.  
Uncertainty represented as 95% HPD intervals widths.
```{r}
mat.mod1 <- mat.mod1 %>% select(!Altimeter)
mat.mod2 <- mat.mod2 %>% select(!Altimeter)
mat.mod1$Unc <-  mat.mod1$hpd_upr - mat.mod1$hpd_lwr
mat.mod1$altimeter <- "barometer only"
mat.mod2$Unc <-  mat.mod2$hpd_upr - mat.mod2$hpd_lwr
mat.mod2$altimeter <- "barometer & laser"

mod.all <- rbind(mat.mod1, mat.mod2)

mod.all$altimeter <- ordered(mod.all$altimeter, levels = c("barometer only", "barometer & laser" )) 

# Summary stats of Uncertainty (HPD interval widths)
mod.all %>% group_by(altimeter) %>% summarise(min = round(min(Unc),2), 
                                              max = round(max(Unc),2), 
                                              mean = round(mean(Unc),2),
                                              sd = round(sd(Unc), 2))


## Plot HPD interval diff w/ altitude
ggplot(mod.all, aes(x = BaroAlt, y = Unc, col = as.factor(altimeter))) + geom_point(size = 2) + 
  labs( x = "altitude (m)", y = "95% HPD interval width (m)") + 
  scale_x_continuous(breaks = seq(10,120, by = 10)) + 
  scale_color_manual(values=c("#F8766D", "#529EFF")) +
  theme_light() + 
  theme(legend.title = element_blank(), legend.position = c(0.8, 0.85)) 
ggsave(file.path('..', 'figures', 'Fig6.jpg'), width = 7, height = 7)
```


# Supplementary Material

# Point-estimates
Comparison 
```{r}
mod.all$altimeter <- ordered(as.factor(mod.all$altimeter), c("barometer only", "barometer & laser")) 

ggplot(mod.all, aes(x = AID, y = TL, ymin = hpd_lwr, ymax = hpd_upr, color = Aircraft)) + 
   geom_pointrange(alpha = 0.75, shape = 4) + geom_point(size = 3.5, shape = 4) +
  xlab("") + 
  ylab('predicted length (m)') +
  facet_wrap(~altimeter) +
  theme(legend.position = "right", axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_abline(aes(intercept=11.2, slope=0), col = "red", lty = 2, alpha = 0.75)  # Cutoff for mature humpback whales (Christiansen et al., 2016)

ggsave(file.path('..', 'figures', 'Supp_FigS2.jpg'), width = 9, height = 7)

```


